---
title: "Spread Calculation"
author: "Hadi"
date: "2023-10-24"
output: html_document
---
Load libraries
```{r}
library(splines2)
library(splines)
library(mgcv)
library (RQuantLib)
library(tidyr)
library(xts)
library(plotly)
library(lubridate)
library (readxl)
library(ggplot2)
library(dplyr)
library(scatterplot3d)
library(tidyverse)
library(plot3D)
library(rgl)
library (zoo)
library(plotly)
library(writexl)
library(YieldCurve)
library(patchwork)
library(gridExtra)
library(stargazer)
library(scales)
library(pdfetch)
library(xtable)
```



Importing Dataset
```{r}

features <- 
  read_excel(path =".../Features.xlsx", na = c ("NA", "#N/A N/A", "#N/A Field Not Applicable", "#N/A Invalid Security"))

Yield <- read_excel(path =".../Yields.xlsx")


```


Longitudinal yields
```{r}

#convert to longitudinal data format
Yield <- Yield %>%  gather(key = "Date", value = "yield", -ID_CUSIP)

#convert data format
Yield$Date <- as.Date(as.numeric(Yield$Date), origin = "1899-12-30")

#removing yields nulls
Yield <- Yield[complete.cases(Yield$yield), ]

#merge yield and feature dataset

#merging Features and Yields with ID_CUSIP
NMuni <- merge(Yield, features, by = "ID_CUSIP")

#Labeling th missing state code
NMuni$`State Code`[is.na(NMuni$`State Code`)] <- "AK"

```

Save dataset
```{r}

saveRDS(NMuni, ".../NMuni.rds" )

```


Subset California dataset based on coupon type

```{r}
California <- 
  NMuni %>% filter(`State Code` == "CA", 
                   CPN_TYP == c ("FIXED", "FIXED, OID", "ZERO COUPON, OID"))
```

Calculate Year fraction at day t
```{r}
for (i in 1:nrow(California)){
  California$YF[i] <- yearFraction(ymd(California$Date[i]) 
                                   , ymd(California$MATURITY[i]), dayCounters = 12)
  
}

```


Add PAR_AMT column to the data set
```{r}
California$PAR_AMT <- 100

California <- California %>% filter(!(YF == 0))
```


remove the outliers based on the z scores (Fig B.1)
```{r}
California$Z_score <- scale(California$yield)
California <- California %>% filter(abs(Z_score)< 3)
```

Save California dataset
```{r}
saverds(California, ".../California.rds")

```



Function for calculating PVs using the observed yield to maturity for approach 1
year_fraction = tau_n
```{r}
bond_price_ytm_a1 <-
function(ytm, year_fraction, coupon,
         face_value, coupon_frequency, bond_type)
{
  
  bond_price <- 0
  n <- year_fraction*coupon_frequency
  if (bond_type %in% c("FIXED", "FIXED, OID") ){
    C <- face_value * coupon / coupon_frequency
    if (n > 0) {
      for (t in (seq(ifelse(n%%1 == 0.00, 1, n%%1), n, by = 1))/coupon_frequency){
        discounted_cashflows <- C / ((1 + ytm ) ^ t)
        bond_price <- bond_price + discounted_cashflows
      }
      final_cashflow <- (face_value) / ((1 + ytm ) ^ (n/coupon_frequency))
      bond_price <- bond_price + final_cashflow
    } else {
      final_cashflow <- face_value + C  
      bond_price <- final_cashflow
    }
  }
  else if (bond_type == "ZERO COUPON, OID"){
    n <- year_fraction
    bond_price <- face_value / ((1 + ytm) ^ n)
  }
  
  else {
    stop("Unsupported")
  }
  
  return(bond_price)
}
  


```


Calculate Green PVs based on the observed YTM 
```{r}
for (i in 1:nrow(California)){
  
  ytm <- California$yield[i]/100
  year_fraction <- California$YF[i]
  coupon <- as.numeric(California$CPN[i])/100
  face_value <- as.numeric(California$PAR_AMT[i])
  coupon_frequency <- as.numeric(California$CPN_FREQ[i])
  bond_type <-California$CPN_TYP[i]
  price_ytm <- 
    bond_price_ytm_a1 (ytm = ytm, coupon = coupon,
                                     face_value = face_value, year_fraction = year_fraction,
                    coupon_frequency = coupon_frequency,
                                     bond_type = bond_type)
  California$GPV_1[i] <- price_ytm
  
} 

```


Function for calculating zero coupon yields
```{r}
zcb_eq_ytm_a1 <- function(zcb_ytm, year_fraction,
                           face_value, bond_type, bond_price, 
                          coupon, coupon_frequency) {
  C <- face_value * coupon / coupon_frequency
  if (bond_type %in% c("FIXED", "FIXED, OID", "ZERO COUPON, OID")) {
    return(((1 + zcb_ytm) ^ year_fraction) - ((face_value + C) / bond_price))
  } else {
    stop("Unsupported bond_type")
  }
}




```

Calculate Equivalent ZCB YTM of green bonds in approach 1 for dataset
```{r}
for (i in 1:nrow(California)){
   
  year_fraction <- California$YF[i]
  bond_price <- as.numeric(California$GPV_1[i])
  face_value <- as.numeric(California$PAR_AMT[i])
  bond_type <- California$CPN_TYP[i]
  coupon_frequency <- as.numeric(California$CPN_FREQ[i])
  coupon <- as.numeric(California$CPN[i])/100
  result <- uniroot(zcb_eq_ytm_a1, interval = c(-1, 20),
                     year_fraction= year_fraction, bond_type = bond_type,
                    face_value= face_value, bond_price= bond_price, coupon = coupon, coupon_frequency = coupon_frequency)
  
  California$G_ZCB_YTM_1 [i] <- round(result$root, digits = 4)*100
  
}

```


Calculate  Reference Rates
Import US Treasury Par yield Curve
```{r}
par_data <- read.csv(".../daily_par_yield-curve-rates-1990-2024.csv")

names(par_data)[-1] <- c("0.083", "0.167", "0.231", "0.333", "0.500", "1", "2", "3",
"5", "7", "10", "20", "30")

par_data$Date <- as.Date(par_data$Date)

par_data_mat <- as.matrix(par_data)
par_data_mat <- par_data_mat[, -1]

par_tby <- as.xts(par_data_mat, order.by = par_data$Date)

par_tby <- na.approx(par_tby, na.rm = FALSE)
```



3D plot for US Treasury Par Yield Curve Rates (Fig C.1)
```{r}

par_tby %>%
  data.matrix() %>% 
  t() %>%
  plot_ly(
    x=as.Date(index(par_tby)),
    y=as.numeric(colnames(par_tby)),
    z=.,
    type="surface"
  ) %>%
  plotly::layout(
    title = "Treasury Par Yield Curve Rates",
           scene = list(
               xaxis = list(title = "Date", gridcolor = "rgb(255, 255, 255)",
                            zerolinecolor = "rgb(255, 255, 255)",
                            showbackground = TRUE,
                            backgroundcolor = "rgb(240, 240,240)"),
               yaxis = list(title = "Maturity", gridcolor = "rgb(255, 255, 255)",
                            zerolinecolor = "rgb(255, 255, 255)",
                            showbackground = TRUE,
                            backgroundcolor = "rgb(230, 230,230)"),
               zaxis = list(title = "Yield (%)", gridcolor = "rgb(255, 255, 255)",
                            zerolinecolor = "rgb(255, 255, 255)",
                            showbackground = TRUE,
                            backgroundcolor = "rgb(220, 220,220)")
               
           ))

```


Visualization:
Curve Fitting for a specific day to test the goodness of fit (US Treasury par yield) (Fig E.1)
```{r}
D <- "2023-03-17"
m1 <- 1

terms <-  sort(rep(c(30,20,10,7,6,3,2,1,0.5,0.25,0.1),1))

#layout(matrix(c(1, 2, 3), ncol = 1, byrow = TRUE))

#3_5 Percent Coupon Rate

x_tr <- as.numeric(sort(names(par_tby))) 
                                                
y_tr <-  as.numeric(par_tby[D, sort(names(par_tby))])

cuts <- summary(x_tr)[c(2,3,5)]

spline_fit_tr <- lm(y_tr ~ bs(x_tr, degree = m1, knots = cuts))

pred_tr <- predict(spline_fit_tr, newdata = data.frame(x_tr = terms), se.fit = TRUE)

se_tr <- qnorm((1 + 0.95) / 2) * pred_tr$se.fit


plot(x_tr,y_tr, xlab = "Terms (annualized)", ylab = "Yield (%)", ylim = c(0,2))

 lines(terms, pred_tr$fit, col = "red")
 lines(terms, pred_tr$fit + se_tr, col = "orange", lty = 2)
 lines(terms, pred_tr$fit - se_tr, col = "orange", lty = 2)

# Original legend with smaller text size
legend("bottomright", legend = c("Smoothed Fit", paste(95, "% Confidence Intervals")),
       col = c("red", "orange"), lty = c(1, 2), pch = c(NA, NA), lwd = c(2, 1), cex = 0.7)

#Add vertical dashed lines at specific x values
vertical_lines_x <- cuts 
abline(v = vertical_lines_x, col = "grey", lty = "dashed")
```


Calculate the  equivalent US treasury yield for the green bonds each day
```{r}
for (i in 1:nrow(California)) {
  x <- as.numeric(sort(names(par_tby)))
  y <- as.numeric(par_tby[California$Date[i], 
                                   sort(names(par_tby))])
  cuts <- summary(x)[c(2,3,5)]
  
  tresury_spl_fit <- lm(y ~ bs(x, degree =3, knots = cuts))

  
  tresury_fitted_values <- predict(tresury_spl_fit, newdata = data.frame(x = as.numeric(California$YF[i])),
                           se.fit = FALSE)
  California$tresury_eq_yield[i] <-  tresury_fitted_values
}
```


Calculate PVs based on the yields Treasury yield curve
```{r}
for (i in 1:nrow(California)){
  
  ytm <- California$tresury_eq_yield[i]/100
  year_fraction <- California$YF[i]
  coupon <- as.numeric(California$CPN[i])/100
  face_value <- as.numeric(California$PAR_AMT[i])
  coupon_frequency <- as.numeric(California$CPN_FREQ[i])
  bond_type <- California$CPN_TYP[i]
  price_ref1 <- 
    bond_price_ytm_a1 (ytm = ytm, coupon = coupon,
                                     face_value = face_value,
                    year_fraction = year_fraction, 
                    coupon_frequency = coupon_frequency,
                                     bond_type = bond_type)
  California$RPV_1[i] <- price_ref1
  
} 

```


Calculate Equivalent reference ZCBYTM1 
```{r}
for (i in 1:nrow(California)){
   
  
  year_fraction <- California$YF[i]
  bond_price <- as.numeric(California$RPV_1[i])
  face_value <- as.numeric(California$PAR_AMT[i])
  bond_type <- California$CPN_TYP[i]
  coupon <- as.numeric(California$CPN[i])/100
  coupon_frequency <- as.numeric(California$CPN_FREQ[i])
  result <- uniroot(zcb_eq_ytm_a1, interval = c(-1, 20),
                     year_fraction= year_fraction, bond_type = bond_type,
                    face_value= face_value, bond_price= bond_price, coupon = coupon, coupon_frequency = coupon_frequency)
  California$R_ZCB_YTM_1 [i] <- round(result$root, digits = 4)*100
  
}

```

Calculate Spread 1 and simple differences of the GZYTM and RZYTM
```{r}

California$S1 <- California$G_ZCB_YTM_1 - California$R_ZCB_YTM_1

```



 


